Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-01 measurements.tex
added caption to figures

The authors implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the Raspberry Pi 2 Model B, which
features four ARM Cortex-A7 cores in a Broadcom BCM2836 \ac{SOC} with a 900 MHz
clock. Each core has a 32 KiB L1 data cache and 32 KiB L1 instruction cache. The
cores share a 512 KiB L2 cache. All the measured results, including the baseline
results, were obtained with NEON-enabled code adopted from the FIFI/KODO library
\cite{kodo2011pedersen}. The NEON extension provides an 128-bit \ac{SIMD}
instruction set to the Raspberry Pi 2. Figures~\ref{enc_dec1024},
\ref{enc_dec128} and \ref{enc_dec16} show the encoding and decoding throughput
in MiB per second for different generation sizes ($g$ = 1024, 128, and 16
respectively). The throughput is the rate of generating $g$ coded packets,
divided by the time the encoder took to perform the task. The size of each coded
packet was fixed to $1536$~bytes since that is the typical size of an Ethernet
frame. The blocked operations were performed dividing the matrices in sub-blocks
of 16,32,64,...,1024 symbols, and the figures only show the best results
(usually block sizes of 16 or 32 symbols since with bigger block sizes, the
operands do not fit in L1 cache).

@@

\paragraph{Baseline encoding} The baseline results involve no recording
of the \ac{DAG} and are performed in a \emph{by-the-book} fashion. The encoder
uses only one thread. And the difference between the non-blocked and blocked
encoding is that in the blocked scheme, the matrix multiplications are performed
dividing the matrices in sub-blocks in order to make the algorithm cache
efficient as described in Section~\ref{sub:implementation-multicore}.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{images/2015-04-18_encoding_decoding_1024.eps}
\caption{Encoding and Decoding performance for g = 1024}
\label{enc_dec1024}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{images/2015-04-18_encoding_decoding_128.eps}
\caption{Encoding and Decoding performance for g = 128}
\label{enc_dec128}
\end{figure}

For $g$ = 128, the differences between the baselines operations show that a
blocked algorithm is 15\% faster than the non-blocked variant. Encoding with
four cores is 2.82 times faster than with a single core. Due to the smaller
matrix sizes, the gain by using blocked operations in the baselines is not that
big compared with $g$ = 1024. For the same reason, the matrix inversion is less
expensive. As a consequence, the decoding throughput is 71\% of the encoding
throughput.

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{images/2015-04-18_encoding_decoding_16.eps}
\caption{Encoding and Decoding performance for g = 16}
\label{enc_dec16}
\end{figure}

When $g$ = 16, the gains of blocked operations are negligible compared with the
non-blocked ones. The reason behind this behavior is that all the data fits in
the L1 cache. For the scheduled version, since the problem to solve is so small,
the gain of using four cores is a factor of 1.68 compared with a single core,
and 1.01 compared with two cores. Therefore, there are no practical benefits in
using four cores instead of two.

The differences in throughput, for all generation sizes, between the blocked
baseline and the scheduled measurements are due the time spent resolving the
dependencies and the scheduling overhead. These effects are negligible for big
generation sizes, while considerable for small matrices.

@@

To compare how much slower is the matrix multiplication with the matrix
inversion for different generation sizes, we ran a set of tests. We used a
single core to perform the operations. We changed the generation sizes,
performed matrices multiplications and matrix inversions, and measured the spent
time doing so. We then calculated the ratio $ratio = \frac{T_{mult}}{T_{inv}}$.
Table~\ref{runtimes} summarizes the results. The bigger the matrix size, the
smaller is the calculated ratio. This means that when the problems are bigger,
the decoding throughput decreases compared with the encoding throughput.

>>>>

The authors implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the Raspberry Pi 2 Model B, which
features four ARM Cortex-A7 cores in a Broadcom BCM2836 \ac{SOC} with a 900 MHz
clock. Each core has a 32 KiB L1 data cache and 32 KiB L1 instruction cache. The
cores share a 512 KiB L2 cache. All the measured results, including the baseline
results, were obtained with NEON-enabled code adopted from the FIFI/KODO library
\cite{kodo2011pedersen}. The NEON extension provides an 128-bit \ac{SIMD}
instruction set to the Raspberry Pi 2. Figures~\ref{enc_dec1024},
\ref{enc_dec128} and \ref{enc_dec16} show the encoding and decoding throughput
in MiB per second for different generation sizes ($g$ = 1024, 128, and 16
respectively). The throughput is the rate of generating $g$ coded packets,
divided by the time the encoder took to perform the task. The size of each coded
packet was fixed to $1536$~bytes since that is the typical size of an Ethernet
frame. The blocked operations were performed dividing the matrices in squared
sub-blocks of 16,32,64,...,1024 operands (words in the finite field) in height
and width, and the figures only show the best results (usually block sizes of
16x16 or 32x32 operands since with bigger block sizes, the operands do not fit
in L1 cache).

@@

\paragraph{Baseline encoding} The baseline results involve no recording of the
\ac{DAG} and are performed in a \emph{by-the-book} fashion. The encoder uses
only one thread. And the difference between the non-blocked and blocked encoding
schemes is that in the blocked scheme, the matrix multiplications are performed
dividing the matrices in sub-blocks in order to make the algorithm cache
efficient as described in Section~\ref{sub:implementation-multicore}.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{images/2015-04-18_encoding_decoding_1024.eps}
\caption{Encoding and Decoding performance for g = 1024 \cite{wunderlich2015network}}
\label{enc_dec1024}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{images/2015-04-18_encoding_decoding_128.eps}
\caption{Encoding and Decoding performance for g = 128 \cite{wunderlich2015network}}
\label{enc_dec128}
\end{figure}

For $g$ = 128, the differences between the baselines operations show that a
blocked algorithm is 15\% faster than the non-blocked variant. Encoding with
four cores is 2.82 times faster than with a single core. Due to the smaller
matrix sizes, the gain when using blocked operations in the baselines is not
that significant when compared with $g$ = 1024. For the same reason, the matrix
inversion is less expensive. As a consequence, the decoding throughput is 71\%
of the encoding throughput.

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{images/2015-04-18_encoding_decoding_16.eps}
\caption{Encoding and Decoding performance for g = 16 \cite{wunderlich2015network}}
\label{enc_dec16}
\end{figure}

When $g$ = 16, the gains of blocked operations are negligible compared with the
non-blocked ones. The reason behind this behavior is that all the data fits in
the L1 cache. For the scheduled version, since the problem to solve is so small,
the gain when using four cores is a factor of 1.68 compared with a single core,
and 1.01 compared with two cores. Therefore, there are no practical benefits in
using four cores instead of two.

The differences in throughput, for all generation sizes, between the blocked
baseline and the single threaded scheduled measurements are due the time spent
resolving the dependencies and the scheduling overhead. These effects are
negligible for big generation sizes, while considerable for small matrices.

@@

To compare how much slower is the matrix multiplication with respect to the
matrix inversion for different generation sizes, we ran a set of tests. We used
a single core to perform the operations. We changed the generation sizes,
performed matrices multiplications and matrix inversions, and measured the time
spent doing so. We then calculated the ratio between these two measured times
times $ratio = \frac{T_{mult}}{T_{inv}}$. Table~\ref{runtimes} summarizes the
results. The bigger the matrix size, the smaller is the calculated ratio. This
means that when the problems are bigger, the decoding throughput decreases
compared with the encoding throughput.

### 2016-06-01 testbed.tex
Remove raw kodo and include kodo-rlnc or kodo-cpp. Editing recoding in RLNC

\subsection{System Overview}
\subsection{Installation and Updating}
\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}

@@

\item Clone the Kodo repository in the server by executing: \\
\texttt{git clone git://github.com/steinwurf/kodo.git} in \texttt{\$HOME}. \\

>>>>

\subsection{System Overview}

This section will be used to present our Raspberry Pi testbed, but also how
it is set up and configured.

The testbed is depicted in \ref{fig:testbed_setup}. It consists of twenty
Raspberry Pi 1 model B rev 2 (Rasp\_01 to Rasp\_20) and 
ten Raspberry Pi 2 model b V1.1 (Rasp2\_01 to Rasp2\_10).
The Raspberry Pis are each equipped with an ethernet interface (eth0) that are
assigned the \ac{IP} addresses listed in the figure.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.6\textwidth]{images/testbed_setup.eps}
\caption{Testbed setup}
\label{fig:testbed_setup}
\end{figure}

\subsubsection{Installing Raspbian}

- 
- Download image (wget)
- Update/Modify image (chroot)
- Burn image
- Bang! linux is running
-

We have installed the official Raspbian linux \ac{OS} on all the devices.

\subsection{Installation and Updating}
\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}

@@

\item Clone the Kodo repository in the server by executing: \\
\texttt{git clone git://github.com/steinwurf/kodo.git} in \texttt{\$HOME}. \textbf{Change the repo to kodo-rlnc or kodo-cpp since just raw kodo is going to be depreceated soon} \\
